---
title: "Data Memo"
subtitle: "Data Science 3 with R (STAT 301-3)"
author: "Allie Tong and Alani Cox-Caceres"

format:
  html:
    toc: true
    embed-resources: true
    code-fold: show
    link-external-newwindow: true
    echo: false
    
execute:
  warning: false

from: markdown+emoji 
---

## Dataset Overview
Our [dataset] (https://www.kaggle.com/datasets/lepchenkov/usedcarscatalog) consists of variables that describe cars in the used car market in Belarus on December 2, 2019. There are 38531 observations. Our target variable will be `is_exchangeable`, which is a factor variable that indicates `TRUE` if the used car can be exchanged with another car and `FALSE` if the used car cannot be exchanged with another car. There are 19 predictors, of which 1 is an integer, 5 are characters, 7 are numeric, and 6 are factors.

## Issues
The dataset was relatively clean but there were a couple issues we encountered. On the dataset, there are 10 features labeled as `feature_#`. These features did not have any descriptions, and the author of the dataset indicated that he was not sure what the features represented. So, we decided to take the features out of the dataset. This could harm the performance of our predictions for `is_exchangeable`, as the features could have had high predictor importance towards our outcome variable.

Potential issues we expect to 

## Exploring Target Variable and Predictors
As seen below, our outcome variable does have issues of imbalance. However, this imbalance is not severe. We will deal with the imabalance by splitting the data accordingly. There is no missingness in our outcome variable.
```{r}
library(tidyverse)
cars <- read_rds(file = "data/processed/cars_clean.rds")

ggplot(cars, mapping = aes(x = is_exchangeable)) +
  geom_bar()
```
Overall, there is no missingness in 18 of our 19 predictors. The predictor `engine_capacity` does have very minor issues of missingness, which we can just impute.
```{r}
library(naniar)
library(kableExtra)

miss_var_summary(cars) %>% 
  kbl() %>% 
  kable_styling()
```

## Data Splitting and Resampling Plan
